\documentclass{article}
\usepackage[margin=0.9in]{geometry}
\usepackage{tikz}
\usepackage{fancyvrb}
\usepackage{graphicx}
\usepackage{needspace}

\input{macro.tex}
\input{../tikz/fgz.tikz.tex}

\title{fgz User Guide}
\author{}
\date{}

\raggedbottom
\setlength{\parindent}{0pt}
\newcommand{\fgzAdvancedScale}{0.62}
\newcommand{\fgzAdvancedPanel}[1]{\scalebox{\fgzAdvancedScale}{\input{#1}}}

\begin{document}
\maketitle

In this repo, build once and then run:

\begin{verbatim}
npm run build
npm run fgz2tex -- examples/guide-minimal.fgz
\end{verbatim}

That writes \texttt{examples/guide-minimal.fgz.tex} by default.

To use fgz in your own project, first install it once:

\begin{verbatim}
npm install --save-dev git+https://github.com/dellaert/fgz.git#main
\end{verbatim}

After that, generate TikZ snippets with:

\begin{verbatim}
npx fgz2tex figures/example.fgz
\end{verbatim}

That writes \texttt{figures/example.fgz.tex} by default. The snippets in this guide
assume the paper preamble already includes \verb|\input{fgz.tikz.tex}|.

\section{Progressive Examples}

Every fgz file starts with \texttt{fgz 1}. Positions are given in TikZ units, and the
generated snippets assume the paper preamble already includes \verb|\input{fgz.tikz.tex}|.

\Needspace{18\baselineskip}
\subsection{Minimal Factor Graph}
Start with two variables and a binary factor. Because the factor has no explicit
position, fgz places it halfway between the first two variables.

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.41\textwidth}
    \centering
    \vspace{0.5cm}
    \input{guide-minimal.fgz.tex}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.55\textwidth}
    \textbf{Source: \texttt{guide-minimal.fgz}}
    \VerbatimInput[frame=single,fontsize=\scriptsize]{guide-minimal.fgz}
  \end{minipage}
\end{samepage}

\Needspace{18\baselineskip}
\subsection{Themes and Known Variables}
Observed factor-graph variables use \texttt{known}. Factors can be positioned explicitly,
and the overall look can be changed with a file-level \texttt{theme}.

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.41\textwidth}
    \centering
    \vspace{0.5cm}
    \input{guide-theme-known.fgz.tex}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.55\textwidth}
    \textbf{Source: \texttt{guide-theme-known.fgz}}
    \VerbatimInput[frame=single,fontsize=\scriptsize]{guide-theme-known.fgz}
  \end{minipage}
\end{samepage}

\Needspace{18\baselineskip}
\subsection{Factor Shape}
Factors are circles by default. Add \texttt{shape=square} when you want a square factor
node instead.

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.41\textwidth}
    \centering
    \vspace{0.5cm}
    \input{guide-factor-shape.fgz.tex}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.55\textwidth}
    \textbf{Source: \texttt{guide-factor-shape.fgz}}
    \VerbatimInput[frame=single,fontsize=\scriptsize]{guide-factor-shape.fgz}
  \end{minipage}
\end{samepage}

\Needspace{18\baselineskip}
\subsection{Per-File Node and Factor Size}
The \texttt{style} line tunes sizing for one fgz file at a time. Here it makes the
variables slightly larger and the factor slightly bolder, while also relaxing the label fit.

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.41\textwidth}
    \centering
    \vspace{0.5cm}
    \input{guide-style-size.fgz.tex}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.55\textwidth}
    \textbf{Source: \texttt{guide-style-size.fgz}}
    \VerbatimInput[frame=single,fontsize=\scriptsize]{guide-style-size.fgz}
  \end{minipage}
\end{samepage}

\Needspace{18\baselineskip}
\subsection{Curved Binary Factors}
For binary factors, a small \texttt{offset=(dx,dy)} moves the factor away from the
midpoint and bends the single factor-graph edge through that new factor position.

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.41\textwidth}
    \centering
    \vspace{0.5cm}
    \input{guide-curved-binary-factor.fgz.tex}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.55\textwidth}
    \textbf{Source: \texttt{guide-curved-binary-factor.fgz}}
    \VerbatimInput[frame=single,fontsize=\scriptsize]{guide-curved-binary-factor.fgz}
  \end{minipage}
\end{samepage}

\Needspace{18\baselineskip}
\subsection{Factor-Graph Parents for Nodes}
Mixed diagrams are allowed when factor-graph variables or knowns act as parents of
Bayes-net nodes. Nodes themselves do not participate in factors.

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.41\textwidth}
    \centering
    \vspace{0.5cm}
    \input{guide-fg-parents-bn.fgz.tex}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.55\textwidth}
    \textbf{Source: \texttt{guide-fg-parents-bn.fgz}}
    \VerbatimInput[frame=single,fontsize=\scriptsize]{guide-fg-parents-bn.fgz}
  \end{minipage}
\end{samepage}

\Needspace{18\baselineskip}
\subsection{Known Nodes and Directed Curve Overrides}
Bayes-net declarations support both latent nodes and observed nodes. A directed
\texttt{curve} statement overrides one implied parent-to-child edge.

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.41\textwidth}
    \centering
    \vspace{0.5cm}
    \input{guide-directed-curve.fgz.tex}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.55\textwidth}
    \textbf{Source: \texttt{guide-directed-curve.fgz}}
    \VerbatimInput[frame=single,fontsize=\scriptsize]{guide-directed-curve.fgz}
  \end{minipage}
\end{samepage}

\Needspace{18\baselineskip}
\subsection{Color Accents}
Colors apply to declarations and to their generated edges. This is useful for marking
active variables, measurements, or elimination steps.

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.41\textwidth}
    \centering
    \vspace{0.5cm}
    \input{guide-color-accents.fgz.tex}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.55\textwidth}
    \textbf{Source: \texttt{guide-color-accents.fgz}}
    \VerbatimInput[frame=single,fontsize=\scriptsize]{guide-color-accents.fgz}
  \end{minipage}
\end{samepage}

\Needspace{18\baselineskip}
\subsection{Macros for LaTeX Labels}
Once the structure is clear, macros keep the source readable while emitting richer math
labels. References stay short in the graph definition, while the rendered labels come
from the macro table.

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.41\textwidth}
    \centering
    \vspace{0.5cm}
    \input{guide-macros.fgz.tex}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.55\textwidth}
    \textbf{Source: \texttt{guide-macros.fgz}}
    \VerbatimInput[frame=single,fontsize=\scriptsize]{guide-macros.fgz}
  \end{minipage}
\end{samepage}

\Needspace{22\baselineskip}

\newpage
\subsection{A Colored Chain Model}
This example combines direct LaTeX-style labels, square factors, colors, and higher-arity
factors in a temporal chain. The red factors are positioned explicitly, while the blue and
green ternary factors rely on midpoint placement along their first two variables. The
document-level style line nudges node size, factor size, label padding, and label font for
this one figure without changing the shared TikZ support file.

\vspace{0.5cm}
\begin{samepage}
  \centering
  \scalebox{0.82}{\input{guide-colored-chain.fgz.tex}}

  \medskip
  \noindent\textbf{Source: \texttt{guide-colored-chain.fgz}}
  \VerbatimInput[frame=single,fontsize=\scriptsize]{guide-colored-chain.fgz}
\end{samepage}

\Needspace{20\baselineskip}
\subsection{A Small Gray Bayes Net}
This example uses the same spacing scale as the chain model, but with only directed edges.
Filled circular nodes highlight selected variables, while \texttt{known\_node} gives a square
parent at the top right.

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.41\textwidth}
    \centering
    \vspace{0.4cm}
    \input{guide-gray-bayes-net.fgz.tex}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.55\textwidth}
    \textbf{Source: \texttt{guide-gray-bayes-net.fgz}}
    \VerbatimInput[frame=single,fontsize=\scriptsize]{guide-gray-bayes-net.fgz}
  \end{minipage}
\end{samepage}

\Needspace{20\baselineskip}
\subsection{A Compact Factor Graph}
This one returns to pure factor-graph structure on the same grid. A known variable sits
above the midpoint factor, while square unary factors anchor the left and bottom edges.

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.41\textwidth}
    \centering
    \vspace{0.4cm}
    \input{guide-compact-factor-graph.fgz.tex}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.55\textwidth}
    \textbf{Source: \texttt{guide-compact-factor-graph.fgz}}
    \VerbatimInput[frame=single,fontsize=\scriptsize]{guide-compact-factor-graph.fgz}
  \end{minipage}
\end{samepage}

\Needspace{18\baselineskip}
\subsection{A Small Directed Triangle}
This final compact example uses the same grid again, but only Bayes-net edges: the square
node at the top is a known parent of both variables, and the right variable also parents the
left one.

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.41\textwidth}
    \centering
    \vspace{0.4cm}
    \input{guide-directed-triangle.fgz.tex}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.55\textwidth}
    \textbf{Source: \texttt{guide-directed-triangle.fgz}}
    \VerbatimInput[frame=single,fontsize=\scriptsize]{guide-directed-triangle.fgz}
  \end{minipage}
\end{samepage}

\Needspace{24\baselineskip}
\newpage
\subsection{Dashed and Labeled Bayes-Net Edges}
The \texttt{edge} statement overrides one implied directed Bayes-net edge to add a dashed
style and a small label beside the arrow. Solid unlabeled edges still come directly from the
\texttt{node} and \texttt{known\_node} parent lists.

\vspace{0.5cm}
\begin{samepage}
  \centering
  \scalebox{0.76}{\input{guide-dashed-bayes-net.fgz.tex}}

  \medskip
  \noindent\textbf{Source: \texttt{guide-dashed-bayes-net.fgz}}
  \VerbatimInput[frame=single,fontsize=\scriptsize]{guide-dashed-bayes-net.fgz}
\end{samepage}

\clearpage
\section{Advanced Example}

\subsection{Toy SLAM Elimination Sequence}
This reconstruction starts from the original SLAM factor graph and then uses five more
fgz files to show the elimination sequence: unary and binary factors, color accents,
factor offsets, and the transition from factor-graph structure toward a Bayes net as
variables are eliminated.

\vspace{0.5cm}
\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.48\textwidth}
    \centering
    \fgzAdvancedPanel{slam-original-factor-graph.fgz.tex}

    \smallskip
    \textbf{(a)} Original factor graph.
  \end{minipage}\hfill
  \begin{minipage}[t]{0.48\textwidth}
    \centering
    \fgzAdvancedPanel{slam-eliminate-l1.fgz.tex}

    \smallskip
    \textbf{(b)} Eliminate $l_1$.
  \end{minipage}\hfill
\end{samepage}

\medskip

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.48\textwidth}
    \centering
    \fgzAdvancedPanel{slam-eliminate-l2.fgz.tex}

    \smallskip
    \textbf{(c)} Eliminate $l_2$.
  \end{minipage}\hfill
  \begin{minipage}[t]{0.48\textwidth}
    \centering
    \fgzAdvancedPanel{slam-eliminate-x1.fgz.tex}

    \smallskip
    \textbf{(d)} Eliminate $x_1$.
  \end{minipage}
\end{samepage}

\medskip

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.48\textwidth}
    \centering
    \fgzAdvancedPanel{slam-eliminate-x2.fgz.tex}

    \smallskip
    \textbf{(e)} Eliminate $x_2$.
  \end{minipage}\hfill
  \begin{minipage}[t]{0.48\textwidth}
    \centering
    \fgzAdvancedPanel{slam-eliminate-x3.fgz.tex}

    \smallskip
    \textbf{(f)} Eliminate $x_3$.
  \end{minipage}
\end{samepage}

\Needspace{34\baselineskip}
\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.48\textwidth}
    \textbf{Source: \texttt{slam-original-factor-graph.fgz}}
    \VerbatimInput[frame=single,fontsize=\tiny,baselinestretch=0.9]{slam-original-factor-graph.fgz}
    \textbf{Source: \texttt{slam-eliminate-l2.fgz}}
    \VerbatimInput[frame=single,fontsize=\tiny,baselinestretch=0.9]{slam-eliminate-l2.fgz}
    \textbf{Source: \texttt{slam-eliminate-x2.fgz}}
    \VerbatimInput[frame=single,fontsize=\tiny,baselinestretch=0.9]{slam-eliminate-x2.fgz}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.48\textwidth}
    \textbf{Source: \texttt{slam-eliminate-l1.fgz}}
    \VerbatimInput[frame=single,fontsize=\tiny,baselinestretch=0.9]{slam-eliminate-l1.fgz}
    \textbf{Source: \texttt{slam-eliminate-x1.fgz}}
    \VerbatimInput[frame=single,fontsize=\tiny,baselinestretch=0.9]{slam-eliminate-x1.fgz}
    \textbf{Source: \texttt{slam-eliminate-x3.fgz}}
    \VerbatimInput[frame=single,fontsize=\tiny,baselinestretch=0.9]{slam-eliminate-x3.fgz}
  \end{minipage}
\end{samepage}

\clearpage
\subsection{Textbook Recursive Dynamics Graph}
This larger factor graph reconstructs the recursive-dynamics layout from the supplied
TikZ example using \texttt{theme textbook}. It combines explicit factor placement,
per-node size tuning for long mechanics labels, and a dense multi-row geometry without
leaving the fgz DSL. \textbf{Source: \texttt{guide-textbook-dynamics.fgz}}

\vspace{0.5cm}
\begin{samepage}
  \centering
  \scalebox{0.85}{\input{guide-textbook-dynamics.fgz.tex}}
\end{samepage}

\clearpage
\subsection{Tracking Message-Passing Panels}
These four panels reconstruct a multitarget-tracking factor-graph decomposition and
exercise the newer fgz features used for this family of figures: labeled square
factors, free text, outline boxes, and a plate for repeated structure.

\vspace{0.5cm}
\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.48\textwidth}
    \centering
    \scalebox{0.68}{\input{guide-tracking-temporal-chain.fgz.tex}}

    \smallskip
    \textbf{(a)}
  \end{minipage}\hfill
  \vspace{0.5cm}
  \begin{minipage}[t]{0.48\textwidth}
    \centering
    \scalebox{0.68}{\input{guide-tracking-association-star.fgz.tex}}

    \smallskip
    \textbf{(b)}
  \end{minipage}
\end{samepage}

\medskip

\begin{samepage}
  \noindent
  \begin{minipage}[t]{0.48\textwidth}
    \centering
    \scalebox{0.68}{\input{guide-tracking-bipartite-block.fgz.tex}}

    \smallskip
    \textbf{(c)}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.48\textwidth}
    \centering
    \scalebox{0.68}{\input{guide-tracking-lifted-box.fgz.tex}}

    \smallskip
    \textbf{(d)}
  \end{minipage}
\end{samepage}

\medskip
\textbf{Panels:} (a) temporal chain; (b) association star; (c) bipartite block; (d)
lifted box construction.

\smallskip
\textbf{Sources:}
\begin{minipage}[t]{0.48\textwidth}
  \texttt{guide-tracking-temporal-chain.fgz}\\
  \texttt{guide-tracking-bipartite-block.fgz}
\end{minipage}\hfill
\begin{minipage}[t]{0.48\textwidth}
  \texttt{guide-tracking-association-star.fgz}\\
  \texttt{guide-tracking-lifted-box.fgz}
\end{minipage}

\end{document}
